---
title: "RICHARDS_Skills_Test_3"
author: "Jessica Richards"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1a - Cleaning Faculty Salaries Data

Cleaning the data was a bit of a long process but I was able to elongate the data
so that there was a column for Rank, Salary, Comp and Number.

```{r loading packages, echo=FALSE, include=FALSE}
library(tidyverse)
library(janitor)
library(broom)
```


```{r cleaning faculty salaries, echo=TRUE}
dat <- read.csv('FacultySalaries_1995.csv')

dat_keep <- dat %>%
  select(-c(AvgProfSalaryAll, AvgProfCompAll, NumInstructors, NumFacultyAll))

dat_both <- dat_keep %>%
  select(-c(starts_with('Num')))

dat_salary <- dat_both %>%
  select(-c(ends_with('Comp'))) %>%
rename('Full' = 'AvgFullProfSalary',
       'Assoc' = 'AvgAssocProfSalary',
       'Assist' = 'AvgAssistProfSalary') %>%
  pivot_longer(c('Full', 'Assoc', 'Assist'), names_to = 'Rank', values_to = 'Salary')

dat_comp <- dat_both %>%
  select(-c(ends_with('Salary'))) %>%
rename('Full' = 'AvgFullProfComp',
       'Assoc' = 'AvgAssocProfComp',
       'Assist' = 'AvgAssistProfComp') %>%
  pivot_longer(c('Full', 'Assoc', 'Assist'), names_to = 'Rank', values_to = 'Comp')

dat_num <- dat_keep %>%
  select(-c(starts_with('Avg'))) %>%
   rename('Full' = 'NumFullProfs',
         'Assoc' = 'NumAssocProfs',
         'Assist' = 'NumAssistProfs') %>%
  pivot_longer(c('Full', 'Assoc', 'Assist'), names_to = 'Rank', values_to = 'Number')
 
dat_clean <- dat_num %>%
  full_join(dat_comp) %>%
  full_join(dat_salary) %>%
  filter(Tier != 'VIIB')


```

This is a sneak peak of the cleaned Faculty Salaries Data. 
```{r head faculty salaries, include=TRUE}
head(dat_clean)



```


# 1b Graphing Faculty Salaries Data

The graphs are separated by Tier and show how average salaries of professors
are affected by both Rank and Tier. 

Based on the graph, Tier I professors on average make the most while Tier IIB 
professors make the least. Full professors in all tiers make the most while Assistant
professors in all tiers make the least. 
```{r graph faculty salaries, echo=FALSE}
dat_clean %>%
  ggplot(aes(x = Rank,
             y = Salary,
             fill = Rank)) +
  geom_boxplot(outlier.color = 'black', outlier.size = 3, size = .8) +
  facet_wrap(~Tier) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45),
        axis.title = element_text(size = 20),
        axis.text = element_text(size = 15),
        title = element_text(size = 20),
        strip.text = element_text(size = 15),
        legend.text = element_text(size = 15))
```


# 2 Creating an Anova Model of Faculty Salaries

Based on the anova summary, State, Tier and Rank all significantly affect the
average salary of professors.
```{r anova model, echo=TRUE, include=TRUE}
anova_model <- aov(data = dat_clean, Salary ~ State + Tier + Rank)
summary(anova_model)

```

# 3 Cleaning Juniper Oils Data

To clean this dataset I tidied the names using clean_names in the janitor package,
then replaced any underscores with hyphens to match the chemical name set provided 
in the instructions. I then elongated the data so that all of the chemical names 
were under chemical_id, and the concentrations were under concentration. Finally
I once again used clean_names to change hyphens in column names to underscores. 
```{r cleaning juniper oils, echo=TRUE, include=TRUE}
dat_oils <- read.csv('Juniper_Oils.csv')

dat_oils <- dat_oils %>%
  clean_names() %>%
  rename_with(~ gsub("_", "-", .))

dat_oils_clean <- dat_oils %>%
  pivot_longer( cols = c("alpha-pinene","para-cymene","alpha-terpineol",
                         "cedr-9-ene","alpha-cedrene","beta-cedrene",
                         "cis-thujopsene","alpha-himachalene","beta-chamigrene",
                         "cuparene","compound-1","alpha-chamigrene","widdrol",
                         "cedrol","beta-acorenol","alpha-acorenol",
                         "gamma-eudesmol","beta-eudesmol","alpha-eudesmol",
                         "cedr-8-en-13-ol","cedr-8-en-15-ol","compound-2",
                         "thujopsenal"), 
                names_to = 'ChemicalID',
                values_to = 'Concentration') %>%
  clean_names()


```


This is a sneak peak of the cleaned Juniper Oils Data.

Since there are so many columns it doesn't display well, but the most important
things are that there are now columns for chemical_id and concentration, 
instead of a ton of columns for each individual chemical. 
```{r head Juniper Oils, echo=FALSE, include=TRUE}
head(dat_oils_clean)

```

# 4 Graphing Juniper Oils Data

```{r graphing Juniper Oils, echo=FALSE, include=TRUE}
dat_oils_clean %>%
ggplot(aes(x = years_since_burn,
           y = concentration)) +
  geom_smooth() +
  facet_wrap(~chemical_id, scales = 'free') +
  theme_minimal()

```

# 5 Modeling Juniper Oils Data

The dataframe tidy_model shows which chemicals' concentrations are significantly 
(p < 0.05) affected by years_since_burn. 
```{r Juniper Oils model, echo=TRUE, include=TRUE}
model <- glm(data = dat_oils_clean, concentration ~ years_since_burn * chemical_id)
tidy_model <- tidy(model) %>%
  filter(p.value < 0.05)

head(tidy_model)

```